---
title: "ISLR Chpater 3 Labs"
output: html_notebook
---
This is an R Notebook containing the chapter 3 labs from 'An Introduction to Statistical Learning with Applications in R'

```{r setup, include=FALSE}
install.packages("MASS")
install.packages("ISLR")
install.packages("car")
library("MASS")
library("ISLR")
library("car")
```

## Boston Data Details
Housing Values in Suburbs of Boston. The Boston data frame has 506 rows and 14 columns.

Columns:

* crim: per capita crime rate by town.
* zn: proportion of residential land zoned for lots over 25,000 sq.ft.
* indus: proportion of non-retail business acres per town.
* chas: Charles River dummy variable (= 1 if tract bounds river; 0 otherwise).
* nox: nitrogen oxides concentration (parts per 10 million).
* rm: average number of rooms per dwelling.
* age: proportion of owner-occupied units built prior to 1940.
* dis: weighted mean of distances to five Boston employment centres.
* rad: index of accessibility to radial highways.
* tax: full-value property-tax rate per \$10,000.
* ptratio: pupil-teacher ratio by town.
* black: 1000(Bk - 0.63)^2 where Bk is the proportion of blacks by town.
* lstat: lower status of the population (percent).
* medv: median value of owner-occupied homes in \$1000s.

## Lab 3.6.2: Simple Linear Regression
### Fit an SLR model to the Data
Fit a simple linear regression model where medv is the response variable and lstat is the predictor varaible

```{r fit-slr, echo=FALSE}
lm.fit=lm(medv~lstat, data=Boston)
# get coefficients
coef(lm.fit)
# get confidence intervals
confint(lm.fit)
# get data from model
summary(lm.fit)
# predict used to produce confidence intervals
predict(lm.fit,data.frame(lstat=c(5,10,15)), interval ="confidence")
# predict used to produce confidence intervals
predict(lm.fit,data.frame(lstat=c(5,10,15)), interval ="prediction")

```

#### Interpretation of the above:

The 95 % confidence interval associated with a lstat value of 10 is (24.47, 25.63), and the 95 % prediction interval is (12.828, 37.28). As expected, the confidence and prediction intervals are centered around the same point (a predicted value of 25.05 for medv when lstat equals 10), but the latter are substantially wider.

### Plot of medv and lstat along with the least sqaures regression line

```{r plot-least-sqr-reg, echo=FALSE}
plot(lstat, medv)
abline(lm.fit)
```

#### Interpretation of the above:

There is some evidence for non-linearity in the relationship between lstat and medv.

### Plot of residuals

```{r plot-residuals, echo=FALSE}
# plot residuals
plot(predict(lm.fit), residuals(lm.fit))
```
### Plot studentized residuals
```{r plot-studentized-res, echo=FALSE}
# plot studentized residuals
plot(predict(lm.fit), rstudent(lm.fit))
```
#### Interpretation of the above:
On the basis of the residual plots, there is some evidence of non-linearity.

### Plot Leverage statistics
```{r plot-leverage-stats, echo=FALSE}
plot(hatvalues (lm.fit))
```
The largest element of the vector, i.e. the largest leverage statistic is:
```{r largest-vect-element, echo=FALSE}
which.max(hatvalues (lm.fit))
```

## Lab 3.6.3: Multiple Linear Regression

### Fit Multiple Linear regression for lstat and age

```{r fit-multi-lin-reg, echo=FALSE}
# fit lstat and age as predictors
lm.fit=lm(medv~lstat+age, data=Boston)
# get a summary of the results
summary(lm.fit)
```


### Fit Multiple Linear Regression for all variables

```{r fit-multi-lin-reg-all, echo=FALSE}
# fit lstat and age as predictors
lm.fit=lm(medv~., data=Boston)
# get a summary of the results
summary(lm.fit)
```
### Fit Multiple Linear Regression for all variables minus age
```{r fit-multi-lin-reg-all-minus-age, echo=FALSE}
# fit lstat and age as predictors
lm.fit=lm(medv~.-age, data=Boston)
# get a summary of the results
summary(lm.fit)
```
Compute variance inflation factors for the above model
```{r compute-variance-inflation-factors, echo=FALSE}
vif(lm.fit)
```

## Lab 3.6.4: Interaction Terms

### Fit a linear regression model for lstat, age and an interaction term (lstat*age)

```{r linear-reg-interaction-term, echo=FALSE}
summary(lm(medv~lstat*age,data=Boston))
```